\large{\bf Looking forwards and backwards through locally regulated populations }

### 2023-04-14 forwards-and-backwards.tex
numbering
one more fig

@@

Simulating these processes and exploring parameter space,
one sooner or later comes upon a strange observation:
with certain parameter combinations, the population spontaneously forms a regular grid
of more or less dicrete patches, separated by areas with nearly no individuals.
The phenomenon is discussed in Section~16.10 of \citet{haller2022evolutionary},
and has been described in similar models, e.g.,
by \citet{britton1990spatial,sasaki1997clumped,hernandezgarcia2004clustering,young2001reproductive,berestycki2009nonlocal}.
Heuristically, this can occur if density-dependent effects of individuals
extend sufficiently farther than the typical dispersal distance:
the patches are stable because new offspring landing between two clumps
effectively compete with \emph{both} neighboring clumps,
while individuals within a clump effectively compete with only one.

@@

\begin{figure}
    \begin{center}
        \includegraphics[width=0.45\textwidth]{figures/clumpy_2d/clumpy_123.density.3.0}
        \comment{TODO} FIGURE
    \end{center}
    \caption{
        Two panels: left is 2D picture of clumped population;
        right is a bumpy expanding wavefront.
        \label{fig:clumping}
    }
\end{figure}

@@

\comment{TODO: the previous definition of $\Pgen^N$ was the following wrong thing; fixup below}
\begin{equation}
\begin{aligned}
    \text{ONLY RIGHT TO $O(1/N)$} \qquad
    \Pgen^N G_f(\eta)
    &:=
    G'(\langle f, \eta \rangle)
        \theta \bigg\langle
        \gamma(x, \eta)
            \int f(z) r(z,\eta) q_\theta(x,dz)
        -
        f(x)\mu_\theta(x, \eta),
        \eta(dx) \bigg\rangle
    \\&\qquad {}
    + \frac{1}{2} G''(\langle f, \eta \rangle)
    \frac{\theta}{N}
    \bigg\langle
        \gamma(x, \eta) \int f^2(y) r(y, \eta) q_\theta(x, dy) + f^2(x) \mu_\theta(x, \eta),
    \eta(dx) \bigg\rangle .
\end{aligned}
\end{equation}

>>>>

\numberwithin{equation}{section}

@@

Simulating these processes and exploring parameter space,
one sooner or later comes upon a strange observation:
with certain parameter combinations, the population spontaneously forms a regular grid
of more or less dicrete patches, separated by areas with nearly no individuals,
as shown in Figure~\ref{fig:clumping}.
The phenomenon is discussed in Section~16.10 of \citet{haller2022evolutionary},
and has been described in similar models, e.g.,
by \citet{britton1990spatial,sasaki1997clumped,hernandezgarcia2004clustering,young2001reproductive,berestycki2009nonlocal}.
Heuristically, this can occur if density-dependent effects of individuals
extend sufficiently farther than the typical dispersal distance:
the patches are stable because new offspring landing between two clumps
effectively compete with \emph{both} neighboring clumps,
while individuals within a clump effectively compete with only one.

@@

\begin{figure}
    \begin{center}
        \includegraphics[width=0.45\textwidth]{figures/clumpy_2d/clumpy_123.density.3.0}
        \includegraphics[width=0.45\textwidth]{figures/clumpy_1d/clumpy_123.steps}
    \end{center}
    \caption{
        \textbf{Left:} A snapshot of individual locations in a two-dimensional simulation
        in which the constant density is unstable
        and a stable, periodic pattern forms.
        \textbf{Right:} Population density in an expanding wave in a one-dimensional simulation
        forming a period pattern;
        each panel shows the wave front in three periods of time;
        within each period of time the wave front at earlier times is shown in blue
        and later times in pink.
        In both cases, $\gamma(m) = 3/(1 + m)$, $\mu \equiv 0.3$, and $r \equiv 1$;
        dispersal is Gaussian with $\sigma=0.2$ and density is measured with $\rho_\gamma(x) = p_9(x)$,
        i.e., using a Gaussian kernel with standard deviation 3.
        \label{fig:clumping}
    }
\end{figure}

@@
